---
title: "mrca clusters"
author: "Stephane"
date: "9/21/2017"
output: html_document
---

```{r}
library(knitr)
read_chunk('test_tmrca_clusters.R')
#opts_chunk$set(dev = c('pdf','png','postscript') )
```

Since we are not using sequence data, our clustering algorithm is based on time dated trees. We added a comparison between the clustering method directly based on patristics distances used in the paper and a network creation method based on time to the most recent common ancestor. 
The approach is similar Leigh-Brown JID 2011 (except they first down-sampled their data selecting sequences under a pairwise genetic distance threshold plus a minimum bootstrap support), in which *"Two nodes are linked by an edge if the highest posterior density time to the most recent common ancestor (MRCA) of the respective viral sequences in the [time dated] tree is less than or equal to some network depth cutoff value."*

The method goes as follows:

- On dated tree, use `mrca` function from `ape` package to obtain $n \times n$ matrix of nodes corresponding to mrca
- Calculate time to mrca for each pair of tips
- Create a link (assign 1) between pair if both time intervals < `threshold` in years
- Use `component` in `igraph` package to assign tips into clusters (single-linkage algorithm)
- Plot by coloring branches

```{r function}
```

First, test on random tree
```{r libs, warning=FALSE, message=FALSE}
```

```{r rnd dated tree}
```

Run
```{r on rnd tree}
```
Works

Now, run on big sim tree. I load the first simulated tree
```{r load sim}
```

I compute and store the matrix of MRCAs and test different thresholds of time to MRCA
```{r big}
```

To change list of clusters into dataframe with size of clusters
```{r helpers}
```

Load results from `hivclustering`
```{r load hivclustering}
```

I compare distribution of cluster sizes at different thresholds of evolution rate / time to mrca.
Great ! Distributions look similar for rate threshold = 0.015 subs/site vs time threshold = 5 years (Sergei Pond did say that somewhere ?)
```{r distribution cluster size}
```

Plot cluster sizes from both methods
```{r scatter}
```
Correlation is `r corcoef`