---
title: "Risk: cluster size vs outdegrees"
author: "Stephane"
date: "29 June 2016"
output:
  html_document:
    fig_width: 6
    fig_height: 6
---

```{r}
library(knitr)
read_chunk('ext-stats_sims.R')
opts_chunk$set(dev = c('png','postscript') )
```

```{r, libs, warning = FALSE}
```
* There is a 10-fold difference in transmission rates between risk level 1 and 2
* In baseline scenario, test sensitivity to detect difference between risk groups

Load tables
```{r, load data}
```

Detect difference in transmission rate by risk level
Univariate
p-values of univariate test: size ~ risk

```{r, u test}
```

Multivariate

```{r, lm}
```
* Note log transformation and no zero values for outdegrees

#### Proportion of tests correctly detecting difference in transmission rates
```{r, results='asis'}
knitr::kable(tab)
```



#### p-values univariate
```{r, boxplot risk 1, fig.path = './figure/', warning = FALSE}
```

```{r, boxplot base risk 1, fig.path = './figure/', warning = FALSE}
```

* SA always detects difference
* neighborhood size and uscd cluster require low threshold to start seeing the difference

#### p-values multivariate
```{r, boxplot risk 2, warning = FALSE}
```

```{r, boxplot base risk 2, warning = FALSE}
```

* Important loss of sensitivity to detection when adjusting
* Need to test scenario with different levels of difference in transmission rates ?

```{r, long}
```

#### Outdegree/sizes vs risk level
```{r, boxplot 3}
```

* note: only use 10 bootstraps and outlier values are censored at 5%
* When difference detected, effect size always underestimated

#### Logistic regression of cluster membership
```{r logistic regression}

```

```{r, results='asis'}
knitr::kable(tab_logit)
```

#### Some stats
Proportion in ucsd clusters by risk
```{r, stats cluster 1}
```

Mean outdegree or size, by risk
```{r, stats cluster 2}
```

Ratio of means (risk level 2 / risk level 1)
```{r, stats cluster 3}
```





