---
title: "Age: Assortativity and associations to cluster size and infector probability"
author: "Stephane"
date: "13 June 2016"
output:
  html_document:
    fig_width: 6
    fig_height: 6
---

```{r}
library(knitr)
read_chunk('assort_matrix.R')
read_chunk('age_regressions.R')
```

```{r include}
```

* I compare SA with both UCSD cluster and _distance neighborhood_ at 4 thresholds
* Here are mostly figures (detail_knitr = `r detail_knitr`)

### Assortativity

* Assortativity matrices for each sims in `age` folder
    + analyzed in `assort_matrix.R`
* UCSD cluster size, neighborhood size, outdegrees, indegrees in `cw_Baseline0.rds` file
* Age matrices computed on HPC in `data/simulations2/age/`

```{r libs, include = detail_knitr}
```

```{r load data, echo = FALSE}
```

```{r aggregate, include = detail_knitr}
```

```{r utility functions, include = detail_knitr}
```

```{r apply assort matrix, include = detail_knitr}
```

##### Assortativity of infector probs by age
```{r heatmap 1, echo = detail_knitr, fig.height = 3, fig.width = 3}
```

##### Assortativity of neighborhood size by age by distance threshold
```{r heatmap 2, echo = detail_knitr}
```

##### Assortativity of cluster size by age by distance threshold
```{r heatmap 3, echo = detail_knitr}
```

* Assortativity mostly detected at younger age
* And at lower threshold

```{r apply assort coef globally, echo = detail_knitr}
```

```{r true assortativity from model0, include = detail_knitr}
```

```{r estimated coefs, include = detail_knitr}
```

```{r boxplot of assort coef, include = detail_knitr}
```

##### Distribution of estimated assortativity coefficient by method
```{r boxplot 1}
```

##### Distribution of estimated assortativity coefficient (log) by method
```{r boxplot 2}
```

* As Erik found, all methods underestimate true assortativity
* At lowest threshold, age assortativity into cluster is closer to truth than for SA but less precisely estimated  

### Regressions

```{r associations, include = detail_knitr}
```

```{r winsorize outliers, include = detail_knitr}
```

```{r reshape, include = detail_knitr}
```

```{r effect winsorize, include = detail_knitr}
```

##### Distribution of outdegrees and cluster sizes at distance threshold = 0.015 by age category
```{r boxplot sizes vs age, echo = detail_knitr}
```

* Data are from 5 bootstrap values
* Outliers have been censored (_winsorized_) to 95% range of values

```{r regressions, include = detail_knitr}
```

##### Regressing age without stage interaction
```{r m1}
```

* Significant associations of size and age
    + At threshold 0.015, UCSD cluster would find `r round(m1[[2]][[2]][4]*100)`%  
    + SA would find `r round(m1[[2]][[2]][1]*100)`%

##### Regressing age by cluster size with stage interaction
```{r m2}
```

##### Regressing age by neighborhood size with stage interaction
```{r m3}
```

* Significant associations of size and age (when adjusting for stage)
    + At threshold 0.015, UCSD cluster would find `r round(m2[[2]][[2]][4]*100)`%  
    + At threshold 0.015, neighborhood would find `r round(m3[[2]][[2]][4]*100)`% 
    + SA would find `r round(m3[[2]][[2]][1]*100)`%
