---
title: "Neigborhood vs ucsd cluster"
author: "Stephane"
date: "4 July 2016"
output: html_document
---

Attempt to understand:
  1. how UCSD cluster are calculated (single or complete linkage)
  2. how neighborhood size differs from cluster size
  
```{r, message = FALSE}
##- understanding difference neighbourhood and ucsd cluster
# rm(list=ls())
library(igraph)
threshold <- '0.015'
####---- path sims ----
path.sims <- 'data/simulations2/model0-simulate'
path.results <- 'data/sim_ucsd_results2'
####---- scenario ----
scenario <- c("Baseline0", "EqualStage0")
scenario <- setNames(scenario, scenario) # useful to name list in lapply
####--- list of distances files ---
list.dist <- lapply(scenario, function(x){
  list.files('RData', full.names = TRUE, 
             path = paste(path.sims, x, '-distances', sep = '') )
})
```

**Distance matrix from first baseline simulations**
```{r}
##- load distances
ldist <- list.dist[["Baseline0"]]
load(ldist[1])
dd <- as.matrix(t(D)) # columns = from, to, distance 
name.sim <- regmatches(ldist[1], regexpr("[0-9]{3,}", ldist[1]))
```

**Choose one cluster as an example**
```{r}
##- load ucsd assignement
##- keep first sim and 0.015 only
cl_Baseline0 <- readRDS( file = paste(path.results, 'list.hivclust.sim.Baseline0.rds', sep = '/') )
# names(cl_Baseline0)
cl <- cl_Baseline0[[threshold]][[name.sim]]
# str(cl)
##- size of cluster
freqClust <- as.data.frame(table(cl$ClusterID), stringsAsFactors = FALSE)
head(freqClust, 10)
```

```{r, include = FALSE}
##- load computed cluster and nbhsize
l_Baseline0 <- readRDS(file = paste(path.results, 'list.sim.ucsd.Baseline0.rds', sep = '/') )
names(l_Baseline0)
l <- l_Baseline0[[threshold]][[name.sim]]
str(l)
```

**Subset of pairwise distances from a given cluster**
```{r}
#- what are the distances within a given cluster ?
seq <- cl[cl$ClusterID == 8,]$SequenceID # 8 # 934 # 1703
el <- dd[dd[,1] %in% seq & dd[,2] %in% seq, ]
el
```

* Cluster #8 is chosen, it has a size 7
* The distances between its members can exceed the threshold (`r as.numeric(threshold)`) which corresponds to connected components or *single-linkage method* (agglomerative until min{d(x,y)} < threshold) 

**Graph showing only distances within threshold**
```{r}
##- restrict edges
el_g <- el[el[,3] < as.numeric(threshold),]
#- make igraph
g1 <- graph.data.frame( as.data.frame(el_g), directed = F )
plot(g1, vertex.shape = 'none')
```

**How neighbourhood size is calculated ?**

For now the first following calculation is performed: for each element of the donor vector, sum distances less than threshold.
```{r}
##- For now 
t1  <- tapply(el[,3], el[,1], function(x) sum(x < 0.015)) # out
t1
```

What probably should be done is summing neighborhood from vertex as donor and recipient
```{r}
##- must add neigborhood for donor AND recipient
t2  <- tapply(el[,3], el[,2], function(x) sum(x < 0.015)) # in 
t <- tapply(c(t1, t2), names(c(t1, t2)), function(x) sum(x) ) # = cluster size if complete linkage method ?
t
```

Now for all individuals.
In my pipeline, neighborhood size is calculated only for first column of distance matrix. And 0 size are added for tree tips not in the matrix 
```{r}
##- how neighbourhood size is calculated ?
##- redo what is done in sim_cluster
all_nbh  <- tapply(dd[,3], dd[,1], function(x) sum(x < 0.015))
# head(all_nbh)
##- neighborhood = 0 if not in distance matrix
nbh_size <- c(all_nbh, rep(0, 12164 - length(all_nbh) ))
# summary(all_nbh)
summary(nbh_size)
##- which is the same as
summary(l$nbhsize)
```

What I should do: 
```{r}
##- now if adding 
all_donor  <- tapply(dd[,3], dd[,1], function(x) sum(x < 0.015))
all_recip  <- tapply(dd[,3], dd[,2], function(x) sum(x < 0.015))
# head(all_donor[order(names(all_donor))])
# head(all_recip[order(names(all_recip))])
all_nbh2 <- tapply(c(all_donor, all_recip), names(c(all_donor, all_recip)), sum) 
summary(all_nbh2)
##- add 0 for other tips
nbh_size2 <- c(all_nbh2, rep(0, 12164 - length(all_nbh2) ))
summary(nbh_size2)
```

**In Erik's code**
```{r}
##- check Erik's calculation
load('Erik_code/age-anal0.0.RData')
tab <- fns2agemat_tables[[1]][['tab']]
summary(tab$nbrhoodSize)
```
Looks like my current values of nbhsize

