---
title: "Regressions on bootstrap UK trees - comparison SA vs Cluster"
output: 
  html_document:
    fig_width: 6
    fig_height: 6
author: "Stephane"
date: "14/4/2016"
fontsize: 11pt
geometry: margin=1in
---


* restrict or not to cohort of sampling 
* add explanatory variables
* categorize age and cd4
* run model and tests

```{r echo=FALSE}
library(knitr) 
opts_chunk$set(echo = FALSE, cache=FALSE)

read_chunk("SA_phydynr.R")
```


Load the libraries and model code:
# ```{r}
#source("regressions_uk.R")
#source("SA_phydynr.R")
# ```
### Calculating phydynR 
```{bs tree to W}

```


## Some simple comparisons of infector probabilities by stage, age, risk level
The total number of transmission events within-sample: 
```{r}

```


## Summary

* Out degree is increasing in age and stage, consistent with interpretation of out-degree as proportional to cumulative number of infections
* These patterns need to be characterised in real data
* Similar analyses need to be done using genetic-distance clustering; it will likely show opposite trends with respect to age and stage; sensitivity to detecting a difference in risk level needs to be characterised
* Counterfactual simulations (equal rates) need to be done to assess false-positive inference of clustering and SA
* Variation in transmission rates by stage is difficult to estimate using SA but Erik will continue to look into it
* Similarly, also need to examine assortativity by age; this can be compared to true level of assortativity in simulations, and can characterise false positive rates for clustering & SA 

