---
title: "Analyze simulations cluster size and outdegrees"
author: "Stephane"
date: "23 May 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=FALSE)
```

Load tables
```{r}
##- Load data (only threshold 0.015)
cw_Baseline0 <- readRDS(file = "data/sim_ucsd_results2/list.sim.clus-outdeg.Baseline0.rds" )
cw_EqualStage0 <- readRDS(file = "data/sim_ucsd_results2/list.sim.clus-outdeg.EqualStage0.rds" )
```

## 1. Detect difference in transmission rate by risk level
### 1.1 Univariate
```{r}
##- U tests on size or outdegree
u.test.risk <- function(df, y){
  U <- wilcox.test(df[df$risk == 2, y], 
                   df[df$risk == 1, y], 
                   alternative = "greater") # "two.sided", "less"
  return(U$p.value)
}

##- p-values for test of cluster size in risk 2 vs risk 1
p_cs <- lapply(cw_Baseline0, function (x)  {
  sapply(x, function(df) u.test.risk(df, y = "size") )
 } )
##- Add p-values for outdegree testing
p_sa <- sapply(cw_Baseline0[[1]], function(df) u.test.risk(df, y = "outdegree") )
p_uni <- c(p_cs, "SA"= list(p_sa) )

##- stats of p-values
sapply(p_uni, summary)

##- table of proportion p-values < 0.05
tab <- data.frame("method" = c(paste("Cluster", names(cw_Baseline0)), "SA"),
                 "proportion univariate" = sapply(p_uni, function(x) mean( x < 0.05)),
                 row.names = NULL)

```


### 1.2 Multivariate
```{r}
##- lm model adjusting for stage of infection
y = "outdegree"
lm.risk <- function(df, y){
  model0 <- "scale(log(y)) ~ scale(risk) * as.factor(stage)"
  model <- sub("y", y, model0)
  ##- pvalue for risk parameter
  p <- coef(summary( lm( model , data = df) ))[2, 4]
  return(p)
}
p_mult_cs <- lapply(cw_Baseline0, function (x)  {
  sapply(x, function(df) lm.risk(df, y = "size") )
 } )
p_mult_sa <- sapply(cw_Baseline0[[1]], function(df) lm.risk(df, y = "outdegree") )
p_mult <- c(p_mult_cs, "SA"= list(p_mult_sa) )
sapply(p_mult, summary)

tab <- cbind(tab, "proportion multivariate" = sapply(p_mult, function(x) mean( x < 0.05)))
row.names(tab) <- NULL
```

#### Proportion of tests correctly detecting difference in transmission rates
```{r, results='asis'}
knitr::kable(tab)
```

* Note: restricting to outdegrees > 0 doesn't change anything
* Interpretation: need to test scenario with different levels of difference in transmission rates

```{r}
##- into a dataframe for ggplot boxplot
head(data.frame(sapply(p_uni, c), check.names = F))
### laaaaaaaaaaa------------------------------------------------------
```



## Associations
## Type 1 / 2 error: detect a difference in risk, age, stage ?
## Regressions
## Assortativity
    1. Calculate mixing matrix
    2. Calculate true assortativity coefficient
    3. Compare with estimated coefficient
## Plots
    1. Distribution cluster sizes, out- and indegrees
    2. Network
  
## Next: change scenarios
  * high, low, no age assortativity
  * high, low, no at-risk group




You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
