---
title: "Analyze simulations cluster size and outdegrees"
author: "Stephane"
date: "23 May 2016"
output: html_document
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(cache=FALSE)
```

```{r}
library(knitr)
read_chunk('ext-stats_sims.R')
```

Load tables
```{r load data}
```

## 0. Plot cluster size and outdegree
```{r long}

```
* Cluster sizes and outdegrees have weak correlation

## 1. Detect difference in transmission rate by risk level
### 1.1 Univariate
```{r}

```


### 1.2 Multivariate
```{r}

```

#### Proportion of tests correctly detecting difference in transmission rates
```{r, results='asis'}
# knitr::kable(tab)
```

* Note: restricting to outdegrees > 0 doesn't change anything (there are no zero values ?)
* Interpretation: need to test scenario with different levels of difference in transmission rates

```{r}
##- into a dataframe for ggplot boxplot or violinplot

```



## Associations
## Type 1 / 2 error: detect a difference in risk, age, stage ?
## Regressions
## Assortativity
    1. Calculate mixing matrix
    2. Calculate true assortativity coefficient
    3. Compare with estimated coefficient
## Plots
    1. Distribution cluster sizes, out- and indegrees
    2. Network
  
## Next: change scenarios
  * high, low, no age assortativity
  * high, low, no at-risk group





